### 2008-12-19 talk/ecoop2009/benchmarks.tex
minor fixes, some comments

In section \ref{sec:tlc-features}, we saw that TLC provides most of the
features that usaully make dynamically typed language so slow, such as
\emph{stack-based VM}, \emph{boxed arithmetic} and \emph{dynamic lookup} of
methods and attributes.

@@

\begin{enumerate}
\item By plain interpretation, without any jitting.
\item With the jit enabled: this run includes the time spent by doing the
  compilation itself, plus the time spent by running the produced code.
\item Again with the jit enabled, but this time the compilation has already
  been done, so we are actually measuring how good is the code we produced.
\end{enumerate}

@@

When doing plain interpretation, we need to create and destroy three temporary
objects at each iterations.  By contrast, the code generated by the JIT does
much better.  At the first iteration, the classes of the two operands of the
multiplication are promoted; then, the JIT compiler knows that both are
integers, so it can inline the code to compute the result.  Moreover, it can
\emph{virtualize} all the temporary objects, because they never escape from
the inner loop.  The same remarks apply to the other two operations inside
the loop.

@@

The computation \emph{per se} is trivial, as it calculates either $-n$ or
$1+2...+n-1$, depending on the sign of $n$. The interesting part is the
polymorphic call to \lstinline{accumulate} inside the loop, because the VM has
no way to know in advance which method to call (unless it does flow analysis,
which could be feasible in this case but not in general).  The equivalent C\#
code we wrote uses two classes and a \lstinline{virtual} method call to
implement this behaviour.

However, our generated JIT does not compile the whole function at
once. Instead, it compiles and executes code chunk by chunk, waiting until it
knows enough informations to generate highly efficient code.  In particualr,
at the time when it emits the code for the inner loop it exactly knows the
type of \lstinline{obj}, thus it can remove the overhead of dynamic dispatch
and inline the method call.  Moreover, since \lstinline{obj} never escapes the
function, it is \emph{virtualized} and its field \lstinline{value} is stored
as a local variable.  As a result, the generated code results in a simple loop
doing additions in-place.

>>>>

\anto{We should say somewhere that flexswitches are slow but benchmarks are so
  good because they are not involved in the inner loops}

In section \ref{sec:tlc-properties}, we saw that TLC provides most of the
features that usaully make dynamically typed language so slow, such as
\emph{stack-based interpreter}, \emph{boxed arithmetic} and \emph{dynamic lookup} of
methods and attributes.

@@

\begin{enumerate}
\item By plain interpretation, without any jitting.
\item With the JIT enabled: this run includes the time spent by doing the
  compilation itself, plus the time spent by running the produced code.
\item Again with the JIT enabled, but this time the compilation has already
  been done, so we are actually measuring how good is the code we produced.
\end{enumerate}

@@

When doing plain interpretation, we need to create and destroy three temporary
objects at each iterations.  By contrast, the code generated by the JIT does
much better.  At the first iteration, the classes of the two operands of the
multiplication are promoted; then, the JIT compiler knows that both are
integers, so it can inline the code to compute the result.  Moreover, it can
\emph{virtualize} (see section \ref{sec:virtuals} all the temporary objects, because they never escape from
the inner loop.  The same remarks apply to the other two operations inside
the loop.

@@

The computation \emph{per se} is trivial, as it calculates either $-n$ or
$1+2...+n-1$, depending on the sign of $n$. The interesting part is the
polymorphic call to \lstinline{accumulate} inside the loop, because the interpreter has
no way to know in advance which method to call (unless it does flow analysis,
which could be feasible in this case but not in general).  The equivalent C\#
code we wrote uses two classes and a \lstinline{virtual} method call to
implement this behaviour.

However, our generated JIT does not compile the whole function at
once. Instead, it compiles and executes code chunk by chunk, waiting until it
knows enough informations to generate highly efficient code.  In particualr,
at the time it emits the code for the inner loop it exactly knows the
type of \lstinline{obj}, thus it can remove the overhead of dynamic dispatch
and inline the method call.  Moreover, since \lstinline{obj} never escapes the
function, it is \emph{virtualized} and its field \lstinline{value} is stored
as a local variable.  As a result, the generated code results in a simple loop
doing additions in-place.

### 2008-12-19 talk/ecoop2009/tlc.tex
answer to a comment, insert another one, fix a typo
minor fixes, some comments

Objects represent a collection of named attributes (much like JavaScript or
Self) and named methods.  At creation time, it is necessary to specify the set
of attributes of the object, as well as its methods.  Once the object has been
created, it is not possible to add/remove attributes and methods.

@@

\begin{itemize}
\item \textbf{Stack manipulation}: standard operations to manipulate the
  stack, such as \lstinline{POP}, \lstinline{PUSHARG}, \lstinline{SWAP}, etc.
\item \textbf{Flow control} to do conditional and unconditional jumps.
\cfbolz{what is the bytecode for conditional jumps?}
\item \textbf{Arithmetic}: numerical operations on integers, like
  \lstinline{ADD}, \lstinline{SUB}, etc.
\item \textbf{Comparisons} like \lstinline{EQ}, \lstinline{LT},
  \lstinline{GT}, etc.
\item \textbf{Object-oriented}: operations on objects: \lstinline{NEW},
  \lstinline{GETATTR}, \lstinline{SETATTR}, \lstinline{SEND}.
\item \textbf{List operations}: \lstinline{CONS}, \lstinline{CAR},
  \lstinline{CDR}.
\end{itemize}

@@

\subsection{TLC features}
\label{sec:tlc-features}
\cfbolz{calling this sections "features" is a bit obscure, since it is more
properties of the implementation}

>>>>

Objects represent a collection of named attributes (much like JavaScript or
Self) and named methods.  At creation time, it is necessary to specify the set
of attributes of the object, as well as its methods.  Once the object has been
created, it is possible to call methods and read or write attributes, but not
to add or remove them.

@@

\begin{itemize}
\item \textbf{Stack manipulation}: standard operations to manipulate the
  stack, such as \lstinline{POP}, \lstinline{PUSHARG}, \lstinline{SWAP}, etc.
\item \textbf{Flow control} to do conditional (\lstinline{BR_COND}) and
  unconditional (\lstinline{BR}) jumps.
\item \textbf{Arithmetic}: numerical operations on integers, like
  \lstinline{ADD}, \lstinline{SUB}, etc.
\item \textbf{Comparisons} like \lstinline{EQ}, \lstinline{LT},
  \lstinline{GT}, etc.
\item \textbf{Object-oriented}: operations on objects: \lstinline{NEW},
  \lstinline{GETATTR}, \lstinline{SETATTR}, \lstinline{SEND}.
\item \textbf{List operations}: \lstinline{CONS}, \lstinline{CAR},
  \lstinline{CDR}.
\end{itemize}

@@

\subsection{TLC properties}
\label{sec:tlc-properties}
